---
title: "Check SOI GDL Access Dataset"
format: 
  html:
    self_contained: true
vignette: >
  %\VignetteIndexEntry{Check SOI GDL Access Dataset}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
knitr:
  opts_chunk: 
    collapse: true
    comment: '#>'
---


```{r setup, message=F}
library(GeoLocatoR)
library(tidyverse)
library(lubridate)
library(DT)
```


```{r}
# Root folder of the dataset (typically the Z-drive)
soi_data_directory <- "/Users/rafnuss/Library/CloudStorage/Box-Box/geolocator_data/UNIT_Vogelzug/"

# Load the geolocator (GDL) database from the access file
gdl <- read_gdl(access_file = file.path(soi_data_directory, "database/GDL_Data.accdb"), filter_col = FALSE)
```

```{r}
data_order_file <- read_gdl_access(access_file = file.path(soi_data_directory, "database/GDL_Data.accdb"))

d <- read_gdl_data(data_order_file[1])
o <- read_gdl_orders(data_order_file[2])
```

# GDL_Orders

Read the table `GDL_Orders` and find inconsistencies:

```{r, warning=FALSE}
problems(o) %>% datatable()
```

## Summarize data per Responsible

```{r}
o %>%
  group_by(ResponsibleP3) %>%
  summarize(
    OrderCount = n(),
    TotalNumberOrdered = sum(NumberOrdered)
  ) %>%
  datatable()
```


# GDL_Data

```{r, warning=FALSE}
d <- read_gdl_data("/Users/rafnuss/Library/CloudStorage/Box-Box/geolocator_data/UNIT_Vogelzug/database/GDL_Data.csv")
problems(d) %>% datatable()
```

## OrderName

No orderName:

```{r}
d %>%
  filter(is.na(d$OrderName)) %>%
  datatable()
```

OrderName not present in GDL_Order:

```{r}
d %>%
  group_by(OrderName) %>%
  summarize(
    GDLCount = n()
  ) %>%
  filter(!(OrderName %in% o$OrderName)) %>%
  datatable()
```

## GDL_ID

No GDL_ID

```{r}
d %>%
  filter(is.na(GDL_ID)) %>%
  datatable()
```

Duplicate GDL_ID AND OrderName:

```{r}
d %>%
  group_by(GDL_ID, OrderName) %>%
  filter(n() > 1) %>%
  ungroup() %>%
  arrange(GDL_ID) %>%
  datatable()
```

Duplicate GDL_ID (only):

```{r}
d %>%
  group_by(GDL_ID) %>%
  filter(n() > 1) %>%
  ungroup() %>%
  arrange(GDL_ID) %>%
  datatable()
```

## GDL_Type, hardware, firmware

```{r}
d %>%
  group_by(GDL_Type, FirmwareVersion, HardwareVersion) %>%
  summarize(.groups = "keep", countTag = n()) %>%
  datatable()
```

```{r}
d %>%
  filter(is.na(GDL_Type) | is.na(FirmwareVersion) | is.na(HardwareVersion)) %>%
  datatable()
```

## Species

```{r}
d %>%
  group_by(Species) %>%
  summarize(countTag = n()) %>%
  arrange(countTag) %>%
  datatable()
```

# Actual data

-\> folder and subfolder structure?

## All file extensions

```{r}
tibble(filename = list.files("/Users/rafnuss/Library/CloudStorage/Box-Box/geolocator_data/UNIT_Vogelzug/data/", recursive = TRUE, full.names = TRUE)) %>%
  mutate(ext = tools::file_ext(filename)) %>%
  group_by(ext) %>%
  summarise(count = n()) %>%
  arrange(count) %>%
  datatable()
```

```{r}
for (i in seq_len(nrow(d))) {
  if (is.na(d$GDL_ID[i])) {
    next
  }
  folder <- list.files(glue::glue("./data/{d$OrderName[i]}/"), pattern = d$GDL_ID[i])

  if (length(folder) == 1) {
    d$gdl_id_version <- folder
  } else if (any(grepl(".glf", folder))) {
    d$gdl_id_version <- folder[1]
  } else if (length(folder) > 1) {
    print(folder)
  }
}
```


# Producation database

## Order model

### Order Information

| Variable Name     | Description                                         | Comments                                                                              |
| ----------------- | --------------------------------------------------- | ------------------------------------------------------------------------------------- |
| OrderID           | Unique identifier for the order                     | Integer                                                                               |
| OrderName         | Name or identifier for the specific order           | Code usualy 6 charachter for species scientific name , 2 letter country code , year?? |
| Species           | Main? species of the bird associated with the order |                                                                                       |
| Client            | Client associated with the order                    |                                                                                       |
| ResponsibleP3     | Person responsible for the order                    |                                                                                       |
| ClientCategory    | Category of the client                              | cooperation, own, purchase                                                            |
| GDL_Type          | Type of geolocator tag ordered                      | duplicated with GDL data                                                              |
| NumberOrdered     | Total number of tags ordered                        |                                                                                       |
| OrderStatus       | Current status of the order                         |                                                                                       |
| NumberDelivered   | Total number of tags delivered                      |                                                                                       |
| DeliveryDate      | Expected delivery date for the order                |                                                                                       |
| DateDelivered     | Actual date when the tags were delivered            |                                                                                       |
| UnitPrice         | Price per unit of the tag                           |                                                                                       |
| Currency          | Currency used for the order                         |                                                                                       |
| InvoiceDate       | Date when the invoice was issued                    |                                                                                       |
| RemarksProduction | Additional remarks related to the production        |                                                                                       |
| RemarksOrder      | Additional remarks related to the order             |                                                                                       |

### Software Settings

| Variable Name                 | Description                                     | Comments |
| ----------------------------- | ----------------------------------------------- | -------- |
| IntervalBlue                  | Interval setting for blue light recording       |          |
| LoggingPeriodEnabled          | Indicates if logging period is enabled          |          |
| StartPeriod1                  | Start time for the first logging period         |          |
| IntervalRGC                   | Interval for recording green channel data       |          |
| GreenEnabled                  | Indicates if green channel is enabled           |          |
| RedEnabled                    | Indicates if red channel is enabled             |          |
| ClearEnabled                  | Indicates if clear channel is enabled           |          |
| LightLevelThreshold           | Threshold level for light detection             |          |
| BaseLoggingInterval           | Base interval for logging data                  |          |
| IntervalFactorLight           | Interval factor for light level recording       |          |
| IntervalFactorAirTemperature  | Interval factor for air temperature recording   |          |
| IntervalFactorBodyTemperature | Interval factor for body temperature recording  |          |
| IntervalFactorPressure        | Interval factor for pressure recording          |          |
| IntervalFactorAcceleration    | Interval factor for acceleration recording      |          |
| IntervalFactorMagnetic        | Interval factor for magnetic field recording    |          |
| AccSamplingRate               | Sampling rate for accelerometer data            |          |
| AccXaxisEnabled               | Indicates if X-axis accelerometer is enabled    |          |
| AccYaxisEnabled               | Indicates if Y-axis accelerometer is enabled    |          |
| AccZaxisEnabled               | Indicates if Z-axis accelerometer is enabled    |          |
| DataloggerEnabled             | Indicates if datalogger is enabled              |          |
| RadioEnabled                  | Indicates if radio transmission is enabled      |          |
| Frequencies                   | Frequencies used for data transmission          |          |
| TransmitPower                 | Power level for data transmission               |          |
| WakeUpTrigger                 | Trigger for waking up the device                |          |
| PatternID                     | Identifier for the data transmission pattern    |          |
| PulseWidth                    | Width of the transmission pulse                 |          |
| PulseInterval                 | Interval between transmission pulses            |          |
| PatternInterval               | Interval for the transmission pattern           |          |
| PulsePattern                  | Pattern of the transmission pulses              |          |
| ActiveAfterWakeUp             | Indicates if the device is active after wake up |          |
| TRSchedule                    | Transmission schedule                           |          |
| LoggingSchedule               | Logging schedule                                |          |

### Hardware Characteristics

| Variable Name             | Description                                | Comments |
| ------------------------- | ------------------------------------------ | -------- |
| LWL                       | Unknown                                    |          |
| LWL_Length                | Length of the unknown item                 |          |
| LWL_Diameter              | Diameter of the unknown item               |          |
| Harness                   | Type of harness used                       |          |
| HarnessMaterial           | Material of the harness                    |          |
| HarnessAttachement        | Method of attaching the harness            |          |
| HarnessThickness          | Thickness of the harness                   |          |
| LegHarnessDiameter        | Diameter of the leg harness                |          |
| BreastHarnessDiameterHead | Diameter of the breast harness at the head |          |
| BreastHarnessDiameterTail | Diameter of the breast harness at the tail |          |
| HarnessDescription        | Description of the harness                 |          |

### Priority Settings

| Variable Name  | Description                     | Comments |
| -------------- | ------------------------------- | -------- |
| PrioritySensor | Priority level for sensor data  |          |
| PriorityMemory | Priority level for memory usage |          |

## Data model

### Tag

| Variable Name      | Description                                               | Comments                                                                       |
| ------------------ | --------------------------------------------------------- | ------------------------------------------------------------------------------ |
| OrderName \*       | Identifier for the specific order of the tags             | Could we make this standard (e.g., year-countrycode-site?-tagtype?)            |
| GDL_ID \*          | Unique identifier for the geolocator tag                  | This should absolutely be unique and standard (e.g., ordername-integer)        |
| GDL_Type \*        | Type of geolocator tag (GDL1, GDL2, GDL3a, GDL3pam, uTag) |                                                                                |
| HardwareVersion \* | Hardware version of the tag (e.g., v1.0)                  | This is currently not standardizsed. It is also present in the \*.setting file |
| FirmwareVersion \* | Firmware version of the tag                               | This is currently not standardizsed. It is also present in the \*.setting file |
| PrintManufacturer  | Manufacturer of the tag                                   | Usually empty (`Teltronic AG` or `Hybrid SA`)                                  |
| FinalAssembly      | Final assembly details of the tag                         | Usually empty (`own` or `teltronic`)                                           |

### Software Settings

| Variable Name                 | Description                                    | Comments |
| ----------------------------- | ---------------------------------------------- | -------- |
| LoggingPeriodEnabled          | Indicates if logging period is enabled         |          |
| UTC_StartLog1                 | UTC start time for the first logging period    |          |
| LightLevelThreshold           | Threshold level for light detection            |          |
| BaseLoggingInterval           | Base interval for logging data                 |          |
| IntervalFactorLight           | Interval factor for light level recording      |          |
| IntervalFactorAirTemperature  | Interval factor for air temperature recording  |          |
| IntervalFactorBodyTemperature | Interval factor for body temperature recording |          |
| IntervalFactorPressure        | Interval factor for pressure recording         |          |
| IntervalFactorAcceleration    | Interval factor for acceleration recording     |          |
| IntervalFactorMagnetic        | Interval factor for magnetic field recording   |          |
| Frequency                     | Recording frequency of the tag                 |          |

### Hardware Characteristic

| Variable Name             | Description                                    | Comments |
| ------------------------- | ---------------------------------------------- | -------- |
| TotalWeight               | Total weight of the tag                        |          |
| LWL                       | ??                                             |          |
| LWL_Length                | Length of the ???                              |          |
| LWL_Diameter              | Diameter of the ???                            |          |
| DateBatteryConnected      | Date when the battery was connected to the tag |          |
| VoltageReady              | Voltage level when the tag was ready           |          |
| SleepCurrent              | Current draw in sleep mode                     |          |
| Harness                   | Type of harness used to attach the tag         |          |
| HarnessMaterial           | Material of the harness                        |          |
| HarnessAttachment         | Attachment method of the harness               |          |
| HarnessThickness          | Thickness of the harness                       |          |
| LegHarnessDiameter        | Diameter of the leg harness                    |          |
| BreastHarnessDiameterHead | Diameter of the breast harness at the head     |          |
| BreastHarnessDiameterTail | Diameter of the breast harness at the tail     |          |

### Species

| Variable Name  | Description                               | Comments |
| -------------- | ----------------------------------------- | -------- |
| species        | Species of the bird equipped with the tag |          |
| Species_origin | Origin of the bird species                |          |

### Equipment

| Variable Name     | Description                                       | Comments                                      |
| ----------------- | ------------------------------------------------- | --------------------------------------------- |
| LongitudeAttached | Longitude of the attachment site                  |                                               |
| LatitudeAttached  | Latitude of the attachment site                   |                                               |
| UTC_Attached      | UTC time when the tag was attached                |                                               |
| DateDeparture     | Date of bird's departure from the attachment site | Useful for light calibration and GeoPressure: |

### Retrieval

| Variable Name    | Description                               | Comments |
| ---------------- | ----------------------------------------- | -------- |
| LongitudeRemoved | Longitude where the tag was removed       |          |
| LatitudeRemoved  | Latitude where the tag was removed        |          |
| UTC_Removed      | UTC time when the tag was removed         |          |
| DateArrival      | Date of bird's arrival at the destination |          |

### Technical Retrieval Information

| Variable Name     | Description                                    | Comments |
| ----------------- | ---------------------------------------------- | -------- |
| UTC_StartRTC      | Start time of the Real-Time Clock on the tag   |          |
| UTC_StopRTC       | Stop time of the Real-Time Clock on the tag    |          |
| UTC_StopReference | Reference stop time for data analysis          |          |
| UTC_FirstData     | UTC time of the first data recorded by the tag |          |
| UTC_LastData      | UTC time of the last data recorded by the tag  |          |
| VoltageEnd        | Voltage of the tag at the end of deployment    |          |
| MemoryUsed        | Memory used by the tag                         |          |
| MagCalib          | Magnetic calibration data                      |          |

### Miscellaneous

| Variable Name | Description                    | Comments |
| ------------- | ------------------------------ | -------- |
| Remarks       | Additional remarks or comments |          |

## Folder structure

`data/*OrderName*/*GDLID*_*date*`

| filename                      | Description          | Comments |
| ----------------------------- | -------------------- | -------- |
| `GDLID_*date*.glf`            | light data           |          |
| `GDLID_*date*.acceleration`   | acceleration data    |          |
| `GDLID_*date*.pressure`       | pressure data        |          |
| `GDLID_*date*.AirTemperature` | air temperature data |          |
| `GDLID_*date*.data`           | ??                   |          |
| `GDLID_*date*.settings`       | ??                   |          |
| `GDLID_*date*.log`            | ??                   |          |
| `GDLID_*date*.rep`            | ??                   |          |
| `GDLID_*date*.settings.bin`   | ??                   |          |

## Notes

### Suggestions of improvement

1. Metadata as markdown or word or excel?
2. Cleaning up datafolder:
   1. `data/*OrderName*/*GDL_ID*`
   2. `data/*OrderName*/*GDL_ID*_*date*` for old version only
   3. no zip, pdf, etc file -> move to another folder? e.g., `data/note`
3. Data validation/test: rmarkdown -> html
4. Access is dying: Move to excel?

### Limitations

- More flexible contributors list with role, contact details, etc. (e.g. <https://r-pkgs.org/description.html#sec-description-authors-at-r>)
- Licenses of data (<https://r-pkgs.org/description.html#the-license-field>)
- More context to the project: description, temporal, spatial, taxonomic extend, websites?, project funding, publication?
- We need a unique number. So far, I've assumed that GDL_ID is unique, but it's not! Also, date in folder is not great, there could be a data version on the setting corresponding to the date of creation.
- We should be able to store more information on the animal (ring numner, sex, age, ssp. size etc...)
- How to record information on resightings, both captured without the tag (tag lost), or seen but not captured. or tag damaged?
- We need an easy way to known how much data is available for each tag and the quality of the data (e.g. pressure), and the date extend of the data can be different for each tag.
- Tag effect require control group. Do we want to store this information too?
- Data format are not standard and not explained

### Processing pipeline

Order -> production -> sending -> equipement -> retrieval -> sending back -> extraction -> bundling -> send data.

- Where are which information stored in (1) Order table, (2) GDL table or (3) settings files?
- Also, which table is storing the possibly different information between order, delivered and retrieved?
