---
title: "Read and use DataPackage"
format: 
  html:
    self_contained: true
vignette: >
  %\VignetteIndexEntry{read-and-use-datapackage}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
knitr:
  opts_chunk: 
    collapse: true
    comment: '#>'
---

```{r setup, message=FALSE}
library(GeoLocatoR)
library(tidyverse)
library(GeoPressureR)
```

## Read Geolocator Data Package

A very cool feature of using zenodo is that you can load the data directly from Zenodo without downloading the files! 

```{r}
pkg <- read_gldp("https://zenodo.org/records/13829930/files/datapackage.json")
```

It's also possible to load the file from a local directory,

```{r, eval=F}
pkg <- read_gldp(file.path("~/", "woodlandkingfisher", "datapackage.json"))
```

## Write to GeoPressureTemplate

You can create a [GeoPressureTemplate](https://github.com/Rafnuss/GeoPressureTemplate) folder from the data package using

```{r}
project_dir <- write_geopressuretemplate(pkg, destdir = "~/Desktop/", overwrite = TRUE)
```               

## Read data with GeoPressureR

Analysis of the tag data is best performed with GeoPressureR

```{r}
id <- "16LN"
withr::with_dir(project_dir, {
  tag <- tag_create(id, assert_pressure = FALSE)
})
````
